Adopting agile in the large: motivations, success factors and challenges

### 2014-05-16 journal/article.tex
Summary table: Change resistance

In an organizational transformation working habits and roles need to change.
However, people are usually disinclined to change and some resistance should be
anticipated in any change process. People will not be willing to change unless
there are good reasons, and the change is preceived easy enough. It is difficult
to attain a buy-in for a change, and even organizations that have a flexible
culture will face resistance [P10]. It should be expected that not everyone will
be willing to change, and some employees will never adapt to the change [P6].
During a transformation period objections to change may become a major reason
for loss in time and productivity [P29].

On a team level numoerous reasons for change resistance were given. For
instance, Fecarotta [P14] described the organization of Boeing as risk-averse
and cautious which slows down any change. The organization was also described
having a deeply rooted status-quo because of high employee retention, which
further hindered change [P14]. In some cases it was reported that people worried
about new roles and responsibilities that the agile model might bring [P44]. For
instance, testers were worried that they would need to take on cross functional
tasks, which would be outside their area of expertise [P18]. Yet another reason
for change resistance was the move from own offices to team areas [P22]. People
were also feeling being monitored because of the increased level of interaction
within the team and between project stakeholders [P45]. One study reported the
cultural change required for middle management as the biggest cause of
resistance [P2].

@@

Another common form of change resistance that was reported was skepticism.
Seffernick describes how management did on one hand acknowledge benefits of
agile, but on the other explained away the possibilitiy to apply it due to
contract reasons, the current matrix organization, and other organizational
practices [P46]. In that case the director of development still pushed the agile
transformation forward. Skepticism was often created by other similar
misconceptions, including that agile does not work for complex products [P5,
P36], agile needs to be implemented in a prescriptive by-the-book way [P9], that
frequent meetings will cause overhead [P18], and that agile equals avoiding
governance and working without a plan [P8].

@@

The way the transformation is initiated will also affect how change resistance
will show. In many cases the initiative to change came from management, but when
presented wrong the initiative becomes a mandate that pepole are not receptive
for. Spayd [P49] summarizes this as: ``Organizations do not change merely
because the boss says so, at least not in the way that is intended''. In that
organization the introduction of CMM had been carried out consistently by a
mandate, but the collaborative nature of agile did not mix well with such a
mindset. A top down mandate may also dilute the understanding of the reasons for
starting the transformation, and the understanding of agile development overall
[P2]. O'Connor describes how team members became sceptical towards agile when
implementing mandaetd changes did not bring any visible benefits [P37].

Organizations had also bad experiences if the implementation of the agile model
was prescribed. For instance, doing Scrum based on checklists of practices would
not work [P41]. Benefield described that it was challenging to find a balance
between independence for teams to choose their way and diluting the value and
structure of Scrum [P6]. Giving room for team autonomy is one of the critical
facets of agile, but it must be ensured that teams do not discard practices that
add value and structure.

The direction in which a tension between management and development is built may
also be opposite. In cases where the change was emerging bottom up the
reluctancy to change was not in the team members but management, causing that
significant organizational change could not be raised above the team level. For
instance, Atlas [P3] describes how it would have required executive support to
integrate the product management office and separate quality assurance groups in
the agile process. In this case, when managers owning projects with many teams
were not involved in the transformation, the agile model was restrained from
spreading beyond team level [P3]. Lack of middle management support was also
seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
position to undermine the entire transformation, and may do so if they do not
participate in, and thus understand, the agile method. Agile brings changes to
some management roles [P2], and lack of understanding of agile development will
leave managers feeling left out [P6]. Still one problem is that if management
does not speak out a goal towards using agile, developers feel that the agile
methods may be replaced by something else at any time [P12].

@@

>>>>

In an organizational transformation working habits and roles need to change.
However, people are usually disinclined to change and some resistance should be
anticipated in any change process. People will not be willing to change unless
there are good reasons, and the change is preceived easy enough. It is difficult
to attain a buy-in for a change, and even organizations that have a flexible
culture will face resistance [P10]. It should be expected that not everyone will
be willing to change, even so that some employees will never adapt to the new
model [P6]. During a transformation period objections to change may become a
major reason for loss in time and productivity [P29]. The main types of change
resistance discovered in the primary studies are summarized in table
\ref{table:challenges_changeresistance}.

\begin{table}[b]
    \begin{tabular}{ l l }
        \toprule
        Types of change resistance           &  References  \\
        \midrule
        General resistance to change         &  P14, P18, P22, P44, P45   \\
        Skepticism towards a new model       &  P5, P8, P9, P18, P36, P46 \\
        Top down mandate creates resistance  &  P2, P12, P37, P49 \\
        Management not willing to change     &  P2, P3, P6, P49 \\
        \bottomrule
    \end{tabular}
    \caption{Types of change resistance reported in studies}
    \label{table:challenges_changeresistance}
\end{table}

Numerous reasons were given for why change resistance emerged. For
instance, Fecarotta [P14] described the organization of Boeing as risk-averse
and cautious, which slows down any change. Any change was further hindered by a
deeply rooted status-quo and high employee retention [P14]. In some cases it was
reported that people worried about new roles and responsibilities that the agile
model might bring [P44]. For instance, testers were worried that they would need
to take on cross functional tasks, which would be outside their area of
expertise [P18]. Yet another reason for change resistance was the move from own
offices to team areas [P22]. People were also feeling being monitored because of
the increased level of interaction within the team and between project
stakeholders [P45].

@@

Another common form of change resistance that was reported was skepticism and
distrust in suitability of agile. Seffernick describes how management did on one hand
acknowledge benefits of agile, but on the other explained away the possibilitiy
to apply it due to contract reasons, the current matrix organization, and other
organizational practices [P46]. Skepticism was often created by other similar
misconceptions, including that agile does not work for complex products [P5,
P36], agile needs to be implemented in a prescriptive by-the-book way [P9], that
frequent meetings will cause overhead [P18], and that agile equals avoiding
governance and working without a plan [P8].

@@

The way the transformation is initiated will affect how change resistance will
show. In several cases the change initiative came from management, but when
presented wrong the initiative becomes a mandate that pepole are not receptive
for. Spayd [P49] summarizes this as: ``Organizations do not change merely
because the boss says so, at least not in the way that is intended''. In that
organization the introduction of CMM had been carried out consistently by a
mandate, but the collaborative nature of agile did not mix well with such a
mindset. A top down mandate may also dilute the understanding of the reasons for
starting the transformation, and the understanding of agile development overall
[P2]. O'Connor describes how team members became sceptical towards agile when
implementing mandaetd changes did not bring any visible benefits [P37].
A problem relating to top down mandate is that if management does not speak out
a goal towards using agile, developers may feel that the agile methods may be
replaced by something else at any time [P12].

The direction in which a tension between management and development is built may
also be opposite. In cases where the change was emerging bottom up the
reluctancy to change was not in the team members but management, causing that
significant organizational change could not be raised above the team level. For
instance, Atlas [P3] describes how executive support would have been required to
extend the agile process to involve the product management office and separate
quality assurance groups. In this case, when managers %for multi team projects
were not involved in the transformation, the agile model was restrained from
spreading beyond the level of development teams [P3]. Lack of middle management
support for change and disinclination to change management culture were seen as
some of the most serious problems in the transformation [P2, P49]. Middle
management is in a position to undermine the entire transformation, and may do
so if they do not participate in, and thus understand, the agile method. Agile
brings changes to some management roles [P2], and lack of understanding of agile
development will leave managers feeling left out [P6].

@@

% Where would this belong?
% Organizations had also bad experiences if the implementation of the agile model
% was being prescribed. For instance, implementing Scrum based on checklists of
% practices would not work [P41]. Benefield described that it was challenging to
% find a balance between independence for teams to choose their way and diluting
% the value and structure of Scrum [P6]. Giving room for team autonomy is one of
% the critical facets of agile, but it must be ensured that teams do not discard
% practices that add value and structure.
